NLIZE: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-24 paper/background.tex
add attention figure

\section{Background}
The target audiences of the proposed tool is domain experts who develop learning model for natural language processing tasks. Therefore, certain domain specific background knowledge is required to fully understand and appreciate the technique discussed in this paper. In this section, we first explain the natural language inference (NLI) task and how it fit into the grand challenges in NLP. Then, we examine the common architecture characteristics shared by many state-of-the-art neural linguistic models. Finally, we discuss the role the attention plays in the model and why attention is closely tided to model interpretability.

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The problem definition is to classify the relationship between a premise sentence and a hypothesis sentence. A simple example is illustrated in Table.~\ref{table:NLI}.

\begin{table}[htbp]
\label{table:NLI}
\centering
\caption{An illustration of natural language inference.}
 \begin{tabular}{c | c c c c} 
 \hline
  & sentences & entail & contradict  & neutral \\ [0.5ex] 
 \hline
 premise & Jim ate an apple. &  -  &  -  & - \\ 
 hypothesis & Jim ate a Fuji apple. &   &  & \checkmark \\
 hypothesis & Jim ate fruits. & \checkmark &   &  \\
 hypothesis & Jim ate an banana. &  & \checkmark & \\
 hypothesis & Tom ate an apple &  &  & \checkmark \\
 %\hline
 %premise & Facebook's IPO electrified the general public. &  -  &  -  & - \\ 
 %hypothesis & Facebook went public. & \checkmark   &  &  \\
 %hypothesis & General Electric went public. &  &   & \checkmark \\
 %hypothesis & People ignored Facebook's IPO. &  & \checkmark & \\

@@

However, in the real word the task can be quick challenge especially for learning algorithm. Take the following sentences are an example:  (P) Facebook's IPO electrified the general public. (H1) Facebook went public. (H2) General Electric went public. (H3) People ignored Facebook's IPO.

The prediction must fall in one of three categories: $\{Entailment, Contradiction, Neutral\}$.
$Entailment$ refers to that the textual information of hypothesis is embedded in premise;
$Contradiction$ means the textual of hypothesis opposes that of premise;
and $Neutral$ implies no conclusion can be made on either $Entailment$ or $Contradiction$.
%
Recently ~\cite{BowmanAngeliPotts2015} proposed a large dataset for this task. We will examine
the decomposable attention model~\cite{parikh2016emnlp} on this dataset with perturbation.

The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
deep comprehension of textual information. Major tasks include Natural Language Inference,
Machine Translation, Questions Answering, Summarization, and etc.
Recently, neural networks models have performed strongly on these tasks.

Therefore, making sense and explaining predictions made by deep neural networks is not only
essential for validating and improving the model, but also becoming a necessity with
increasing demands for model accountability (e.g., what is the evidence for making the decision)
and model fairness (e.g., is the prediction affected by the bias in the training data).

@@

>>>>

\section{Background}
The target audiences of the proposed tool is domain experts who develop learning model for natural language processing tasks. Therefore, certain domain specific background knowledge is required to fully understand and appreciate the technique discussed in this paper. In this section, we first explain the natural language inference (NLI) task and how it fit into the grand challenges in NLP. Then, we examine the common architecture characteristics shared by many state-of-the-art neural network models. Finally, we discuss the role the attention plays in the model and why attention is closely tided to model interpretability.

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The problem definition is to classify the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence. 
The prediction must fall in one of three categories: $\{Entailment, Contradiction, Neutral\}$.
$Entailment$ refers to that the textual information of hypothesis is embedded in premise;
$Contradiction$ means the textual of hypothesis opposes that of premise;
and $Neutral$ implies no conclusion can be made on either $Entailment$ or $Contradiction$.
A simple example is illustrated in Table.~\ref{table:NLI}.

\begin{table}[htbp]
\label{table:NLI}
\centering
\caption{An illustration of natural language inference.}
 \begin{tabular}{c | c c c c} 
 \hline
  P / H & sentences & entail & contradict  & neutral \\ [0.5ex] 
 \hline
 premise & Jim ate an apple. &  -  &  -  & - \\ 
 hypothesis & Jim ate a Fuji apple. &   &  & \checkmark \\
 hypothesis & Jim ate fruits. & \checkmark &   &  \\
 hypothesis & Jim ate an banana. &  & \checkmark & \\
 hypothesis & Tom ate an apple &  &  & \checkmark \\
 %\hline
 %premise & Facebook's IPO electrified the general public. &  -  &  -  & - \\ 
 %hypothesis & Facebook went public. & \checkmark   &  &  \\
 %hypothesis & General Electric went public. &  &   & \checkmark \\
 %hypothesis & People ignored Facebook's IPO. &  & \checkmark & \\

@@

However, in real word, the task can be quite challenging, especially for learning algorithms. Take the following sentences as an example (here \textbf{P} refer to premise, \textbf{H} refer to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public. (\textbf{H1}) Facebook went public. (\textbf{H2}) General Electric went public. (\textbf{H3}) People ignored Facebook's IPO. The literal semantic similarity between ``Electric'' and ``electrified'' will like trick the model to predict \textbf{H2} as \emph{entailment}. The model likely will also not be able to understand the link between ``went public'' and ``IPO'', therefore, mistake \textbf{H1} as \emph{neutral}.

\shusen{How does the NLI related to the other challenges in NLP}
The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
deep comprehension of textual information. Major tasks include Natural Language Inference,
Machine Translation, Questions Answering, Summarization, and etc.
Recently, neural networks models have performed strongly on these tasks.
%
Therefore, making sense and explaining predictions made by deep neural networks is not only
essential for validating and improving the model, but also becoming a necessity with
increasing demands for model accountability (e.g., what is the evidence for making the decision)
and model fairness (e.g., is the prediction affected by the bias in the training data).
%
Recently ~\cite{BowmanAngeliPotts2015} proposed a large dataset for this task. We will examine
the decomposable attention model~\cite{parikh2016emnlp} on this dataset with perturbation.

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.9\linewidth]{attentionIllustration}
 \caption{Attention illustrate. The soft alignment between two sentences.}
\label{fig:attention}
\end{figure}
